---
#This script plots Tomistoma schlegelii presence records sourced from; both published and unpublished datasets, 211 from Stuebing et al. (2006), 113 from Sideleau (2023), 36 from Dr Agata Staniewicz and 23 from The Global Biodiversity Information Facility (2023), with additional data from; museum specimen records, sightings, published journal articles and archival field expedition journals#
---
#Load packages#
```{r }
  library(sf)
  library(dplyr)
  library(patchwork)
  library(tidyverse)
  library(sp)
  library(rnaturalearth)
  library(ggplot2)
  
```
  
  
  
#Load and wrangle data#
```{r }
  
data<-read.csv("C:/Users/student/OneDrive/Conservation, Evol, Ecol/Project Two/Tomistoma Distributions Reproducible Code/Tomistoma Distributions Reproducible Code/Tomistoma schlegelli locality data .csv")
```

#Limit columns#
```{r }
  data<-dplyr::select(data, 'Lat', 'Long', 'Extent', 'Date', 'Keep')
```


#Filter NAs out#
```{r }
  data <- data %>%
    filter(trimws(Keep) != "N")
```

#Remove NAs from date
```{r }
  data <- data %>%
    filter(complete.cases(Date))
```


#Order date#
```{r }
  data <- data %>%
    arrange(Date)
```

#Remove rows with values greater than 50 in 'Extent' column, keeping NAs
```{r }
  df_filtered <- data[is.na(data$Extent) | data$Extent <= 50, ]
```


#Print the filtered data frame
```{r }
  data<-print(df_filtered)
```

#Check all columns are numeric#
```{r }
  sapply(data,class)
```
  
#Load range states#
```{r }
range_states<-read.csv('Range states.csv')
select(range_states, 'Lat', 'Long')
  
``` 
  
#Plotting#
#Helper functions for plotting#

```{r }

  remove_y <-
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())
  remove_x <-
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank())
  
  
```  
  
#Convert the Tomistoma coordinates into spatial points then sf format#
```{r }
  
  data <-
    data %>%
    st_as_sf(coords = c("Long", "Lat"), crs = 4326)
```  
  
#Convert the range states coordinates into spatial points then sf format#
```{r }
  
  range_states <-
    range_states %>%
    st_as_sf(coords = c("Long", "Lat"), crs = 4326)
```  
  
  
#Make a base map of the land#
```{r }
baseMap <-
rnaturalearth::ne_countries(returnclass = 'sf')
```   
  
#Plot base map#
```{r }
  ggplot(baseMap) +
    geom_sf(fill = "lightsteelblue1", lwd = 0.4) +
    theme_void()
```   
  

# Choose coordinates to limit to South East Asia#
```{r }
se_asia_bbox <- c(94.816452, -11.046168,
                    154.245815, 25.735938)
xlim_se_asia <- c(se_asia_bbox[1], se_asia_bbox[3])
ylim_se_asia <- c(se_asia_bbox[2], se_asia_bbox[4])
  
``` 
  
#Plot points on map#
```{r }
  
map<-ggplot(baseMap) +
geom_sf(fill = "palegreen", lwd = 0.4) +
# Add species
geom_sf(data = data, colour = "darkgreen", size = 3, alpha=0.5) +
# Add equator line
geom_hline(aes(yintercept = 0), colour = "black") +
geom_sf(data = range_states, colour = "black",shape = 15, size = 1) +
coord_sf(xlim = xlim_se_asia, ylim = ylim_se_asia, expand = TRUE) +
theme_minimal() +
remove_x + # gets rid of x axis as we don't need it
remove_y # gets rid of y axis
  
```
#Here the current distributions are plotted for the intro#

#Read in the IUCN range maps#
```{r }
iucn <- 
  st_read(dsn = "C:/Users/student/OneDrive/Documents/R/CROCODILES_ALLIGATORS") %>%
  st_as_sf() %>%
  st_transform(4326)
```

#Limit IUCN data to Tomistoma schlegelii#
```{r }
iucn2<-subset(iucn, binomial == "Tomistoma schlegelii")
```

#Load basemap#
```{r }
baseMap <-
  rnaturalearth::ne_countries(returnclass = 'sf')
```

#Choose coordinates to limit to South East Asia#
```{r }
se_asia_bbox <- c(94.816452, -11.046168,
                  154.245815, 25.735938)
xlim_se_asia <- c(se_asia_bbox[1], se_asia_bbox[3])
ylim_se_asia <- c(se_asia_bbox[2], se_asia_bbox[4])
```

#Look at the Tomistoma schlegelii distibutions#
```{r }
ggplot(baseMap) +
  geom_sf(fill = "palegreen", lwd = 0.4) +
  theme_void() +
  geom_sf(fill = "darkgreen", colour = "darkgreen",
          data = iucn2, show.legend = FALSE) +
  geom_sf(data = range_states, colour = "black",shape = 15, size = 1) +
  coord_sf(xlim = xlim_se_asia, ylim = ylim_se_asia, expand = TRUE)

```

